---
title: "MTH-235 Exam 3"
author: "Dennis Ulichney"
date: '`r gsub("(\\D)0", "\\1", format(Sys.Date(), "%B %d, %Y"))`'
output: pdf_document
editor_options:
  chunk_output_type: inline
---

# Instructions

* This exam is due **by 11:59 PM on Thursday, May 6** and should be  
submitted in Blackboard (under Assignments / Exams).
* Like always, for **every** question, your task is to do two things:
    1. Write **one line** of code in the answer box that outputs your answer to 
the question at hand.
    2. If you need more than one line of code to answer the question, put the 
other lines of code in the "supporting work" box.
* Be sure that you have electronically signed this exam (see below).    
 
# Important Details
* You are **not** allowed to discuss the exam with any other person, living or 
dead.
* You **are** allowed to consult sources from within our Blackboard site, such 
as:
    * The cheat sheet
    * Old Colab worksheets or keys
    * Old assignment submissions or keys
* You **are** allowed to consult documentation within RStudio.
* You are **not** allowed to consult other internet sources besides our own
Blackboard site. In particular, Stack Overflow is off limits for this exam,
including searching the site (even if not posting questions there).
* Control-flow measures like `if` / `ifelse`, `while`, or `for` are fine to use.
* The point of every question is to use a Monte Carlo simulation, even if 
directly calculating the answer using analytic methods is straightforward.
* If you are unsure about whether something is allowed or not, check with me
first.

# Scoring

Problems are listed with their points (not all problems have equal weight).
the bonus problems are, as always, optional. There are a total of 70 points
on this exam.


\vfill

By signing below, you certify that:

* You have read all the guidelines on this sheet and have followed them all.
* You have not discussed the exam with any other person. 
* You have not used any disallowed source.
* Your solutions are all based on your own understanding, and you could recreate
them with no assistance.

Signature:
<!-- Erase the entire next line and replace it with your full name to sign. -->
Dennis Ulichney

\newpage

# Exam

## Problem 1

```{r majordf, include = FALSE}
majordf <- read.csv(
  "https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/all-ages.csv")
```
The `majordf` data frame, which can be loaded into your workspace by running the
above code block, contains information about college majors based on data from 
the 2010-2012 American Community Survey (ACS). 

- Original data: \url{https://www.census.gov/programs-surveys/acs/
microdata.html}
- Cleaned data: \url{https://github.com/fivethirtyeight/data/blob/master/
college-majors/all-ages.csv}
- Corresponding article: \url{https://fivethirtyeight.com/features/
the-economic-guide-to-picking-a-college-major/}

## Problem 1a (5 pts)

In the `majordf` data frame, how many different major categories are there?

### Supporting Work
```{r}

```

### Answer
```{r}
length(unique(majordf$Major_category))
```

(Correct answer: 16.)

---

## Problem 1b (10 pts)

Continuing with `majordf`: What is the name of the major that has the highest 
number of `Employed` graduates on record?

### Supporting Work
```{r}

```

### Answer
```{r}
majordf$Major[majordf$Employed == max(majordf$Employed)]
```

(Correct answer not disclosed for this problem, but: remember that you can view
the data frame inside RStudio and click variable names to sort by columns. This
is a pretty easy way to see for yourself what the correct answer should be. Your
task for this problem is, of course, to write code that generates the name of a
particular major.)

---

## Problem 1c (10 pts)

Continuing with `majordf`: Consider only the collection of majors that have at 
least 70,000 employees and a median salary of \$50,000 or more. What is the name
of the major in this collection with the highest unemployement rate?

### Supporting Work
```{r}
validmajors <- majordf$Major[majordf$Employed >= 70000 & majordf$Median >= 50000]
validrates <- majordf$Unemployment_rate[majordf$Employed >= 70000 & majordf$Median >= 50000]
```

### Answer
```{r}
validmajors[validrates == max(validrates)]
```

(Correct answer: Again, not shown -- but with the sorting tools available in the
RStudio IDE you can, and should, verify this for yourself.)

---

## Problem 1d (5 pts)
Continuing with `majordf`: there is an apparent problem in the data frame.
Specifically, there is one major for which the reported number of `Employed`
graduates is *less* than the reported number of `Employed Full Time Year Round`
graduates. What is the name of this major? (Write code that outputs the name.)

### Supporting Work
```{r}

```

### Answer
```{r}
majordf$Major[majordf$Employed < majordf$Employed_full_time_year_round]
```

(Correct answer not disclosed for this problem.)

---

## Problem 2 

```{r, include = FALSE}
# don't edit this block
alc_sea <- c(0.073, 0.104, 0.152, 0.118, 0.090, 0.161)
alc_high <- c(0.130, 0.168, 0.146, 0.144, 0.103, 0.141)
```

Some have theorized that the effect of alcohol consumption on the body is much 
greater at higher altitudes. To test this theory, a scientist randomly selected 
12 subjects and divided them into two groups of 6 each. One group was 
transported to an altitude of 12,000 feet, and each member in the group ingested
100 cubic centimeters (cm$^3$) of alcohol. The members of the second group were 
taken to sea level and given the same amount of alcohol. After two hours, the 
amount of alcohol in the blood of each subject was measured (measurements in 
grams per 100 cm$^3$). The data are given in the `alc_sea` and `alc_high` 
vectors in the block above.

To run a hypothesis test, our null hypothesis will be that there is no true
difference between the groups; our alternative hypothesis will be that the
median of the `alc_high` alcohol measurements are greater than those from the
`alc_sea` group. 

## Problem 2a (5 pts)

In a classical statistics class (i.e. one not based on Monte Carlo statistics),
a student might execute something called the Mann-Whitney-Wilcoxon test (also
known as the Mann-Whitney U-test), which detects differences in medians between
two groups. Run the code block below to see the use of such a test in R.

```{r}
# don't edit this block
wilcox.test(alc_high, alc_sea, alternative = "g")  # Mann-Whitney-Wilcoxon test
```

Your question: What does that output imply about the original research question
(i.e. "...the effect of alcohol consumption on the body is much greater at
higher altitudes")? Your answer needs to clearly explain the conclusion that one
should draw from the hypothesis test above.

<!-- Type your answer after the > -->

> Running this code gives me a p-value of 0.197, which is high enough that the
null hypothesis cannot be disproven.

(Correct answer not disclosed for this problem.)

---

## Problem 2b (5 pts)

The output of the code block above gave a $p$-value. Which of these is the 
correct interpretation of that $p$-value? 

### Answer

<!-- answer by converting one [ ] to [X] -->

- [ ] The probability that there **is** a difference between the groups is $0.197.$
- [ ] The probability that there is **no** difference between the groups is $0.197.$
- [ ] Assuming that there **is** a difference between the groups, the probability of
having a gap between samples as large as the one observed is $0.197$.
- [X] Assuming that there is **no** difference between the groups, the probability of
having a gap between samples as large as the one observed is $0.197$.

(Correct answer not disclosed for this problem.)

---

## Problem 3 
A study examined expressions of commitment between partners in heterosexual 
romantic relationships. The study specifically wanted to investigate whether
there is a difference between males and females in terms of which person was the
first in the relationship to say, "I love you."

In the survey, 40 couples were surveyed. Of those 40, 28 responded that the male
in the relationship had said "I love you" before the female in the relationship
did.

## Problem 3a (5 pts)

In words, what was their null hypothesis ($H_0$)? Your answer should be in non-
technical language that someone who had never taken a statistics class would
understand.

### Answer

<!-- Type your answer after the > -->

> Since the scientists are looking to see if there is a difference between
males and females, their null hypothesis would be that there is no difference
between the two.

(Correct answer not disclosed for this problem.)

---

## Problem 3b (5 pts)

Should a one-sided alternative hypothesis be used in this case, or a two-sided
alternative hypothesis? Explain your reasoning.

### Answer

<!-- Type your answer after the > -->

> A two-sided alternative hypothesis should be used because the scientists are
simply looking to see if there is a difference between males and females, not
whether or not one says they love the other person first more or less frequently.

(Correct answer not disclosed for this problem.)

---

## Problem 3c (5 pts)

Next, **visualize** the study by placing it in the context of many simulations
under the null hypothesis. 

### Supporting Work
```{r}
Males_First_Out_of_40_Couples <- replicate(10000, sum(sample(0:1, 40, replace = TRUE)))
```

### Answer
```{r}
hist(Males_First_Out_of_40_Couples)
abline(v = 28, col = "blue")
```

(Correct answer: a histogram with a vertical line to distinguish to results of
the particular study. It's fine if you want to skip specification of the
`breaks` here.)

---

## Problem 3d (5 pts)

Next, **evaluate** the extremity of the sample by computing the $p$-value.

### Supporting Work
```{r}

```

### Answer
```{r}
mean(Males_First_Out_of_40_Couples >= 28)
```

(Correct answer not disclosed for this problem.)

---

## Problem 3e (5 pts)

Finally, **interpret** your work above. Does the study indicate a difference
between males and females in terms of who says "I love you" first? Explain
how you know.

### Answer

<!-- Type your answer after the > -->

> Yes, the study does indicate a difference between males and females. I can say
this because the p-value of the original experiment's result, assuming the
null hypothesis is true, is rather small, only being about .0085. 

(Correct answer not disclosed for this problem.)

---

## Problem 4 (5 pts)

This question is borrowed from Francis Su, a professor of Mathematics at Harvey
Mudd University (see [\underline{this excellent blog post}](https://www.francissu.com/post/7-exam-questions-for-a-pandemic-or-any-other-time)).

Reflect on the mathematical virtue of **persistence**. Consider one homework 
problem you have worked on this semester that you struggled to understand and 
solve, and explain how the struggle itself was valuable. In the context of this
question, describe the struggle and how you overcame the struggle. You might 
also discuss whether struggling built aspects of character in you (e.g. 
endurance, self-confidence, competence to solve new problems) and how these 
virtues might benefit you in later ventures.

<!-- Type your answer after the > -->
> Although its really more of a type of problem than a single, specific one, I had
a considerable amount of trouble understanding how to get the determinant of a
matrix larger than 2x2 in my Linear Algebra class using cofactor expansion.
This was largely because the examples in the notes for the class used different
rows or columns for the operation with no discernible pattern. It wasn't until
after I talked to the teacher that I learned that I can simply use any row or
column that I want. Since then, I always use the top row of the matrix just
because I like to keep things consistent. When it comes to character-building,
I feel that this case didn't have a large impact because, at this point, I am
already a very dedicated student. I do think, however, that this helped teach
me that the solutions to my problems are not always as complicated as I think
they are.

(Please be as specific as you are able to be.)

---

## Bonus Problem 1 (2 pts)

Household incomes in the US are **not** normally distributed; that is, a 
histogram of US household incomes would not have a bell curve shape, but would 
instead show a rightward skew because of the small number of earners whose 
incomes vastly outpace the majority of incomes. 

For an approximation, we can assume that the mean US household income is 50
and the true standard deviation of US household incomes is 20, both measured
in thousands of dollars. If you take a random sample of 28 households, what is
the probability that the average household income will exceed \$60K?

Remember: incomes are **not** normally distributed, and I didn't tell you how
they **are** distributed.

### Supporting Work
```{r}

```

### Answer
```{r}
mean(replicate(10000, rnorm(28, 50, 20) > 60))
```

(Correct answer not disclosed for this problem.)

## Bonus Problem 2 (3 pts)

A person is standing on a large two-dimensional space. He spins around in a 
random angle and takes a step 1 foot in that direction. After making that step,
he spins around again in a different random angle and takes a second step
(irrespective of where he was after his first step). He does this for a total of
ten steps. What is the probability that, after his tenth step, he is within a
1-foot radius of his starting position? (Some geometry is required to solve this
problem.)

### Supporting Work
```{r}
tenstep <- function(){
  angles <- runif(10, 0, 2*pi)
  x <- 0
  y <- 0
  for (angle in angles) {
    x <- x + cos(angle)
    y <- y + sin(angle)
  }
  sqrt(x**2 + y**2)
}
```

### Answer
```{r}
mean(replicate(10000, tenstep() <= 1))
```

(Correct answer not disclosed for this problem.)

<!-- Self-assessment for exams seems weird, so let's skip it. -->

