---
title: "MTH-235 Exam 1"
author: "Dennis Ulichney"
date: '`r gsub("(\\D)0", "\\1", format(Sys.Date(), "%B %d, %Y"))`'
output: pdf_document
editor_options:
  chunk_output_type: inline
---

# Instructions

* This exam is due **by 11:59 PM on Thursday, February 25** and should be 
submitted in Blackboard (under Assignments / Exams).
* Like always, for **every** question, your task is to do two things:
    1. Write **one line** of code in the answer box that outputs your answer to 
the question at hand.
    2. If you need more than one line of code to answer the question, put the 
other lines of code in the "supporting work" box.
* Be sure that you have electronically signed this exam (see below).    
 
# Important Details
* You are **not** allowed to discuss the exam with any other person, living or 
dead.
* You **are** allowed to consult sources from within our Blackboard site, such 
as:
    * The cheat sheet
    * Old Colab worksheets or keys
    * Old assignment submissions or keys
* You **are** allowed to consult documentation within RStudio.
* You are **not** allowed to consult other internet sources besides our own
Blackboard site. In particular, Stack Overflow is off limits for this exam,
including searching the site (even if not posting questions there).
* Control-flow measures like `if` / `ifelse`, `while`, or `for` are disallowed
for this exam.
* The point of every question is to use a Monte Carlo simulation, even if 
directly calculating the answer using analytic methods is straightforward.
* If you are unsure about whether something is allowed or not, check with me
first.

# Scoring

Each non-bonus problem is worth ten points; partial credit will be considered.
The bonus problem is optional and is worth 5 points. There are a total of 70 
points on this exam.


\vfill

By signing below, you certify that:

* You have read all the guidelines on this sheet and have followed them all.
* You have not discussed the exam with any other person. 
* You have not used any disallowed source.
* Your solutions are all based on your own understanding, and you could recreate
them with no assistance.

Signature:
<!-- Erase the entire next line and replace it with your full name to sign. -->
Dennis Ulichney

\newpage

# Exam

## Problem 1

If you roll ten dice (D6), what is the probability that between two and four 
(inclusively) of them will land on the number 6?

### Supporting Work
```{r}
rollcount <- function(){
  sixcount <- sum(sample(1:6,10,replace=T) == 6)
  sixcount >= 2 & sixcount <= 4
}
```

### Answer
```{r}
mean(replicate(10000,rollcount()))
```

(Correct answer: $\frac{3,359,375}{6,718,464}$, which is very close to -- but
not exactly equal to -- 50\%.)

---

## Problem 2

Suppose you roll ten dice (D6); make a histogram representing the **number of 
distinct values** you see would see from those ten rolls. 

Example: If you rolled `3, 4, 6, 6, 5, 5, 2, 5, 6, 2`, then this should appear
in the histogram as a 5 because you saw five distinct values (all possible 
values except 1).

Another example: if you rolled `4, 4, 4, 4, 4, 4, 4, 4, 4`, then this should
appear in the histogram as a 1 because you saw just one distinct value.

### Supporting Work
```{r}
tenroll <- function(){
  length(unique(sample(1:6,10,replace=T)))
}
```

### Answer
```{r}
hist(replicate(10000,tenroll()),breaks=0.5:6.5)
```

(Correct answer: a histogram with bars of varying heights ranging from 1 to 6;
for full credit, the bars should have no gaps between them and each bar should
correspond clearly to exactly one whole number.)

---

## Problem 3

Two people are playing a competitive board game. Player A is allowed to roll two
D6's and declare her score to be the larger of the two dice; player B instead
rolls three D10's and declares his score to be the average of the three dice. In
the event of a tie, Player A is declared the winner. What is the probability 
that Player B wins?

<!-- To reiterate: Player A rolls 2 x D6, while Player B rolls 3 x D10. -->

### Supporting Work
```{r}
gameroll <- function(){
  rollA <- max(sample(1:6,2,replace=T))
  rollB <- mean(sample(1:10,3,replace=T))
  rollB > rollA
}
```

### Answer
```{r}
mean(replicate(10000,gameroll()))
```

(Correct answer not disclosed for this problem.)

---

## Problem 4

An airline finds that if a person books a ticket on a certain flight, they have 
a $95\%$ chance of showing up for the flight. If the airline sells 160 tickets 
for a flight with only 155 seats, what is the probability that a seat will be 
available for every person holding a reservation and planning to fly?

### Supporting Work
```{r}
seatcheck <- function(){
  sum(sample(0:1,160,replace=T,prob=c(5,95))) <= 155
}
```

### Answer
```{r}
mean(replicate(10000,seatcheck()))
```

(Correct answer not disclosed for this problem.)

---

## Problem 5

If you draw cards from a standard deck of 52 without replacement, what is the
expected number of draws required to obtain your first heart? (Your task for
this question is to use the `cards` vector below.)

```{r cards, include = FALSE}
cards <- 
  paste0(c("A", "2", "3", "4", "5", "6", "7", "8", "9", "T", "J", "Q", "K"),
         rep(c("H", "C", "S", "D"), each = 13))
  # format of cards is: [value][suit], such as "3H" for 3 of hearts,
  # "7C" for 7 of clubs, "TD" for 10 of diamonds
```

### Supporting Work
```{r}
drawH <- function(){
  min(which(substr(sample(cards),2,2) == "H"))
}
```

### Answer
```{r}
mean(replicate(10000,drawH()))
```

(Correct answer not disclosed for this problem.)

---

## Problem 6

In the game Bingo, a dealer draws several balls from a bin. There are 75 balls
in the bin, each labeled with both a letter and a number. The labels are B-1, 
B-2, ..., B-15, I-16, I-17, ..., I-30, N-31, N-32, ..., N-45, G-46, G-47, ...,
G-60, O-61, O-62, ..., O-75. (Only the letters B, I, N, G, O are used, and each
appears with 15 different numbers.) If you draw three balls without replacement 
from such a bin, what is the probability that they all have the same letter? 

There is no preloaded vector that you are required to use for this problem; 
construct whatever you need to answer the question.

### Supporting Work
```{r}
BINGO <- paste(rep(c("B","I","N","G","O"),each=15),c(1:75),sep="-")
lettermatch <- function(){
  length(unique(substr(sample(BINGO,3,replace=F),1,1))) == 1
}
```

### Answer
```{r}
mean(replicate(10000,lettermatch()))
```

(Correct answer: $\frac{91}{2701} \approx 0.0337$)

---

## Problem 7

A student answers a multiple-choice examination question that offers four
possible answers. The probability that the student knows the answer to each
question is $0.8$, and the probability that the student does not know (and
therefore must guess) is $0.2$. If the student guesses, she has a $1$ in $4$
chance of selecting the correct answer. What is the overall probability that the
student will get one particular question right? 

(Recall: no `if` or `ifelse` uses are permitted on this exam.)

### Supporting Work
```{r}
choice <- function(){
  sample(c(1,sample(0:1,1,prob=c(3,1))),1,prob=c(8,2)) == 1
}
```

### Answer
```{r}
mean(replicate(10000,choice()))
```

(Correct answer: $0.85$)

---
## Bonus Problem

Write a function which takes a whole number input $n$ and outputs the answer to
the following question:

> If you take $n$ distinct objects and rearrange them in a random order, what is
the expected number of objects that will be in their same original starting
position?

> Example: If you rearrange the five letters ABCDE, you might obtain BDCAE. In
this example, two letters (C and E) are in their original starting position.

Your **Answer** should be a test of this function on the input $n = 555$. Note
that the function must be fully parametrized in order to receive credit -- that
is, the number $n = 555$ cannot be hard-coded into it.

(To be clear: the Monte Carlo simulation should occur entirely within the
function you define. This is not how we have typically done things, but it is a
perfectly sensible to do.)

### Supporting Work
```{r}
samespot <- function(n){
  sum(sample(1:n) == 1:n)
}
```

### Answer
```{r}
mean(replicate(10000,samespot(555)))
```

(The correct answer is the expected number of the 555 objects that remain in
their original place when rearranged in a random order, but I am not going to
tell you what that number is.)

<!-- Self-assessment for exams seems weird, so let's skip it. -->
