---
title: "MTH-235 Exam 2"
author: "Dennis Ulichney"
date: '`r gsub("(\\D)0", "\\1", format(Sys.Date(), "%B %d, %Y"))`'
output: pdf_document
editor_options:
  chunk_output_type: inline
---

# Instructions

* This exam is due **by 11:59 PM on Tuesday, March 30** and should be 
submitted in Blackboard (under Assignments / Exams).
* Like always, for **every** question, your task is to do two things:
    1. Write **one line** of code in the answer box that outputs your answer to 
the question at hand.
    2. If you need more than one line of code to answer the question, put the 
other lines of code in the "supporting work" box.
* Be sure that you have electronically signed this exam (see below).    
 
# Important Details
* You are **not** allowed to discuss the exam with any other person, living or 
dead.
* You **are** allowed to consult sources from within our Blackboard site, such 
as:
    * The cheat sheet(s)
    * Old Colab worksheets or keys
    * Old assignment submissions or keys
* You **are** allowed to consult documentation within RStudio.
* You are **not** allowed to consult other internet sources besides our own
Blackboard site. In particular, Stack Overflow is off limits for this exam,
including searching the site (even if not posting questions there).
* Control-flow measures like `if` / `ifelse`, `while`, or `for` are fine to use.
* The point of every question is to use a Monte Carlo simulation, even if 
directly calculating the answer using analytic methods is straightforward.
* If you are unsure about whether something is allowed or not, check with me
first.

# Scoring

Each non-bonus problem is worth ten points; partial credit will be considered.
The bonus problem is optional and is worth 5 points. There are a total of 70 
points on this exam.


\vfill

By signing below, you certify that:

* You have read all the guidelines on this sheet and have followed them all.
* You have not discussed the exam with any other person. 
* You have not used any disallowed source.
* Your solutions are all based on your own understanding, and you could recreate
them with no assistance.

Signature:
<!-- Erase the entire next line and replace it with your full name to sign. -->
Dennis Ulichney
\newpage

# Exam

## Problem 1

If you draw five cards from a standard deck of 52 without replacement, what is
the probability that you will draw at least one face card? (A "face card" is a
card that is either a Jack, Queen, or King. Your task for this question is to 
use the `cards` vector below.)

```{r cards, include = FALSE}
cards <- 
  paste0(c("A", "2", "3", "4", "5", "6", "7", "8", "9", "T", "J", "Q", "K"),
         rep(c("H", "C", "S", "D"), each = 13))
  # format of cards is: [value][suit], such as "3H" for 3 of hearts,
  # "7C" for 7 of clubs, "TD" for 10 of diamonds
```


### Supporting Work
```{r}
p1 <- function(){
  sum(substr(sample(cards, 5), 1,1) %in% c("J","Q","K")) >= 1
}
```

### Answer
```{r}
mean(replicate(10000, p1()))
```

(Correct answer: $6221/8330 \approx 0.7468$)

---

## Problem 2

If you roll three dice, what is the probability that at least one of the dice is
a 5 or 6, given that the sum of the dice is ten or less?

### Supporting Work
```{r}
p2 <- function(){
  threeroll <- sample(1:6, 3, replace=T)
  (sum(threeroll == 5 | threeroll == 6) >= 1)[sum(threeroll) <= 10]
}
```

### Answer
```{r}
mean(unlist(replicate(10000, p2())))
```

(Correct answer: $4/9$. I thought about obfuscating this answer, as I did with 
some other ones on this test, but I decided not to. I hope that you find that
this makes the exam a bit friendlier and encourages you to continue doing your
best work.)

---

## Problem 3

The systolic blood pressure in healthy adult humans can be modeled with a normal
distribution with mean 112 mmHg and standard deviation 10 mmHG. If a random
sample of 7 humans is collected, what is the probability that their average
blood pressure is above 110 mmHG?

### Supporting Work
```{r}
p3 <- function(){
  mean(rnorm(7, 112, 10)) > 110
}
```

### Answer
```{r}
mean(replicate(10000, p3()))
```

(Correct answer not disclosed for this problem.) 

---

## Problem 4

On a given work day, a pizza delivery driver needs an amount of gas that is
uniformly distributed between 3 and 5 gallons. The price of gas fluctuates from
day to day and can be modeled as uniformly distributed between \$2 and \$3. What
is the expected cost incurred by the driver?

### Supporting Work
```{r}
p4 <- function(){
  runif(1, 3, 5) * runif(1, 2, 3)
}
```

### Answer
```{r}
mean(replicate(10000, p4()))
```

(Correct answer: $\$10$.)

---

## Problem 5

A bin contains five components from supplier A, three from supplier B, and two 
from supplier C. Suppose someone draws three components without replacement from
this bin. Given that they did *not* draw one component from each of the three
suppliers, what is the probability that they drew all three components from the
same supplier?

### Supporting Work
```{r}
p5 <- function(){
  draw <- sample(rep(c("A", "B", "C"), times = c(5, 3, 2)), 3)
  (draw[1] == draw[2] & draw[2] == draw[3])[!identical(sort(draw), c("A", "B", "C"))]
}
```

### Answer
```{r}
mean(unlist(replicate(10000, p5())))
```

(Correct answer not disclosed for this problem.)

---

## Problem 6

Chris is a Cleveland Browns fan. When the Browns lose a game, he has a 90\%
chance of stress-eating an entire pan of brownies and becoming ill. When the
Browns win a game, during his celebrations he still has a 25\% chance of eating
an entire pan of brownies and becoming ill. Suppose the Browns are given a 40\% 
chance of winning a particular game. Given that Chris becomes ill the following 
weekend, what is the probability that the Browns won?

### Supporting Work
```{r}
p6 <- function(){
  win <- sample(0:1, 1, prob = c(60, 40))
  
  if(win == 1){
      binge <- sample(0:1, 1, prob = c(75, 25))
  }
  
  else {
      binge <- sample(0:1, 1, prob = c(10, 90))
  }
  
  (win == 1)[binge == 1]
}
```

### Answer
```{r}
mean(unlist(replicate(10000, p6())))
```

(Correct answer: $5/32 \approx 0.1563$)

---

## Problem 7

Suppose you roll dice and keep track of a running sum until the first time the
running sum is at least 100. Make a histogram of the final die rolled needed to
get over that threshold. 

### Example

Suppose that after rolling several dice, your running sum is 90. If your next 
few die rolls are 3, 5, 1, 2, then your respective running sums are 93, 98, 99, 
101; the roll of 2 is the final die needed to cross the 100 mark. Your task is
to recreate this process many times and report the value of that final die.

### Supporting Work
```{r}
p7 <- function(){
  sum <- 0
  while(sum < 100) {
    roll <- sample(1:6, 1)
    sum <- sum + roll
  }
  roll
}
```

### Answer
```{r}
hist(replicate(10000, p7()), breaks = .5:6.5)
```

(Correct answer: a histogram with bars of varying heights ranging from 1 to 6;
for full credit, the bars should have no gaps between them and each bar should
correspond clearly to exactly one whole number.)

---

## Bonus Problem

Suppose you draw five cards from a deck without replacement. Given that you draw
a straight flush, what is the probability that you draw a royal flush?
(Definitions of these terms are below.)

### Definitions

* A **flush** is a hand consisting of five cards of the same suit.
* A **straight** is a hand consisting of five cards that have adjacent values;
for instance, a hand containing 34567 together is a straight. Note that J, Q,
and K count as numerical values 11, 12, and 13 respectively for the purposes of
a straight. An ace (A) counts as *either* 1 *or* 14, but not both 
simultaneously; that is, A2345 is a straight, as is TJQKA, but QKA23 is not.
* A **straight flush** is a hand that is both a straight and a flush.
* A **royal flush** is a straight flush with the highest possible values; that
is, it is TJQKA of the same suit.

### Supporting Work
```{r}
# From what I can tell, this solution is mostly complete but fails to remove the A's correctly, likely
# due to an issue with variable scope.
bonus <- function(){
  draw <- sample(c("A", 2:13), 5)
  if(draw[1]=="A"){
    draw[1] <- 1
  }
  else if(draw[5]=="A"){
    draw[5] <- 14
  }
  else if("A" %in% draw){
    draw[which(draw=="A")] <- 0
  }
  (all(draw == 10:14))[(draw[5]-draw[4]==1)&(draw[4]-draw[3]==1)&(draw[3]-draw[2]==1)&(draw[2]-draw[1]==1)]
}
```

### Answer
```{r}
#mean(unlist(replicate(10000, bonus())))
```

(Correct answer not disclosed for this problem.)

<!-- Self-assessment for exams seems weird, so let's skip it. -->

