---
title: "Group Project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Part 1: Finds the difference between the proportion of deaths with the killer nurse and the proportion of deaths without her:
```{r}
proj_3 <- read.delim("C:/Users/Owner/Downloads/proj_3.txt") # Imports data.
wGilbert <- proj_3$Patient[proj_3$GilbertWorked. == "Yes"]
noGilbert <- proj_3$Patient[proj_3$GilbertWorked. == "No"]

OriginalDiff <- mean(wGilbert == "Death") - mean(noGilbert == "Death")
```


## Part 2: Creates a function that shuffles the data and computes a new difference of proportions:
```{r}
shuffleDiff <- function(){
  proj_3$Patient <- sample(proj_3$Patient)
  
  wGilbert2 <- proj_3$Patient[proj_3$GilbertWorked. == "Yes"]
  noGilbert2 <- proj_3$Patient[proj_3$GilbertWorked. == "No"]

  mean(wGilbert2 == "Death") - mean(noGilbert2 == "Death")
}

mean(replicate(10000, shuffleDiff()))
```


## Part 3: Finds the p-value of the original difference of proportions against the distribution of shuffled differences:
```{r}
# Null Hypothesis: Gilbert's presence had no effect.
# Alternative Hypothesis: There were more deaths when Gilbert was around.

mean(replicate(10000, shuffleDiff() >= OriginalDiff))
```


## Part 4: Graphs the data from shuffleDiff and shows where on the graph the original difference of proportions falls:
```{r}
hist(replicate(10000, shuffleDiff()), breaks = seq(-.15, .15, by = .01))
abline(v = OriginalDiff, lty = 3, col = "blue")

# The result of this code shows that the difference of proportions found in the
# original data does not fall on the distribution of the shuffled data, resulting
# in a p-value of 0.
```

