---
title: "MTH-235 Assignment 10"
author: "Dennis Ulichney"
date: '`r gsub("(\\D)0", "\\1", format(Sys.Date(), "%B %d, %Y"))`'
output: pdf_document
editor_options:
  chunk_output_type: inline
---


All assignments and exams will be submitted through R Markdown Notebooks like
these. For **every** question, your task is to do two things:

1. Write **one line** of code in the answer box that outputs your answer to the
question at hand.

2. If you need more than one line of code to answer the question, put the other
lines of code in the "supporting work" box.



## Problem 1a

A company claims that no more than 0.1\% of its products are defective --  but
you suspect the real defective rate is higher. (For the sake of this problem, we
will assume the worst case -- that is, exactly 0.1\% of its products are 
defective.) A random sample of 1,433 products reveals three defective units. Is 
this evidence of a problem? First, **visualize** this sample by placing it in 
the context of what many simulations involving taking 1,433 products with a
0.1\% defective rate as the company claims.

### Supporting Work

```{r}
p1 <- function(){
  sum(sample(0:1, 1433, replace = TRUE, prob = c(99.9, .1)))
}
```

### Answer

```{r}
hist(replicate(10000, p1()), breaks = -.5:10.5)
abline(v = 3, col = 'blue')
```

Correct answer: (A histogram that mostly descends from left to right, with a 
colored vertical line to distinguish the 3. Be sure to set the `breaks`; one 
helpful hint is that the number of defectives is very, very rarely larger than 
10.)

---

## Problem 1b

Reconsider the previous problem. **Evaluate** the likelihood of the sample by
computing the $p$-value -- that is, the probability that three or more units 
would be defective out of 1,433.

### Supporting Work

```{r}
p1 <- function(){
  sum(sample(0:1, 1433, replace = TRUE, prob = c(99.9, .1)))
}
```

### Answer

```{r}
mean(replicate(10000, p1() >= 3))
```

(Correct answer: about $0.1744$)

---

## Problem 1c

Does the sample indicate evidence that the company's claim is wrong? Why or why
not?

### Answer

<!-- new question type: short answer 

Answer by typing text after the quote arrow below ">". Be sure to look at the
knitted PDF to make sure that your answer appears reasonably.

Example:
> This is my answer line
--> 

> The sample is not evidence that the company's claim is wrong, it is
not particularly unusual with the defect rate they have given (p > .1).

(Correct answer not disclosed for this problem.)

---

## Problem 2

You want to know if a coin is fair. You flip it 100 times and it lands on heads
32 of those times. What type of test should you use?

### Answer

<!-- New question type: multiple-choice

Answer by placing replacing the whitespace in a box with an X.
Example: What is the name of the university where you are currently studying?

- [ ] Cleveland State
- [ ] John Carroll
- [X] Baldwin Wallace
- [ ] Ohio Northern

-->

- [ ] a left-tailed test, because the simulated outcomes have a bell curve
shape centered around 50 and 32 is on the left side of them
- [X] a two-sided test, because we didn't have a suspicion **before the sample
was collected** to lean one way or another

(Correct answer not disclosed for this problem.)

---

## Problem 3

If a right-tailed test gives a $p$-value of $0.022$, what would the $p$-value of
the corresponding two-tailed test be?

### Answer

<!-- answer by converting one [ ] to [X] -->

- [ ] $0.011$
- [ ] $0.022$
- [X] $0.044$
- [ ] we need more information to determine the answer

---

## Problem 4a

A recent drug trial conducted by Pfizer aimed to investigate the effectiveness
of its vaccine for the novel coronavirus. The experiment consisted of taking
large numbers of participants, some of whom were given the vaccine, and some of
whom were given a placebo. Then, the number of infected participants among the
two groups was recorded. In words, what was their null hypothesis ($H_0$)? 
(Your answer should be in non-technical language that someone who had never
taken a statistics class would understand.)

### Answer

<!-- Type your answer after the > -->

>  Since Pfizer is trying to prove that their vaccine is effective, the
null hypothesis would be that their vaccine is not effective.

(Correct answer not disclosed for this problem.)

---

## Problem 4b

The $p$-value in the aforementioned experiment was found to be less than
$0.0001$. What does this mean about the effectiveness of the vaccine?

### Answer

<!-- Type your answer after the > -->

> A p-value of .0001 indicates that the vaccine is very much effective.

(Correct answer not disclosed for this problem.)

---

## Problem 5a

Suppose you set up a hypothesis test to investigate the relationship between a
lack of sleep and test scores. You use the following hypotheses:
- null hypothesis, $H_0$: a lack of sleep has no effect on test scores
- alternative hypothesis, $H_a$: a lack of sleep lowers test scores
If you compute a $p$-value of $0.0243$, what does this imply about your research
question? (Give an answer using non-technical language that would make sense to
a person who has never taken a statistics course. Use language specific to this
problem.)

### Answer

<!-- Type your answer after the > -->

> A p-value of .0243 indicates a moderate amount of evidence that a lack of
sleep lowers test scores.

(Correct answer not disclosed for this problem.)

---

## Problem 5b

Consider the same setup as problem 5a. If the $p$-value you computed was instead
equal to $0.1780$, what would this imply about your research question? (As
before, give a non-technical answer.)

### Answer

<!-- Type your answer after the > -->

> A p-value of .1780 indicates that a lack of sleep most likely does not
lower test scores.

(Correct answer not disclosed for this problem.)

---

## Problem 5c

Consider the same setup as problem 5a. If the $p$-value you computed was instead
equal to $0.0003$, what would this imply about your research question? (As
before, give a non-technical answer.)

### Answer

<!-- Type your answer after the > -->

> A p-value of .0003 indicates that a lack of sleep almost certainly lowers
test scores.

(Correct answer not disclosed for this problem.)

---

## Problem 6

You wonder if a six-sided die is fair. You roll it 100 times and it lands on the
six side 23 times. Does this constitute good evidence that it is biased?

Your answer should include *four* required elements: a null hypothesis (in 
text), an alternative hypothesis (in text), a p-value (calculated based on some
simulations), and a conclusion (in text).

### Supporting Work
```{r}
p6 <- function(){
  sum(sample(1:6, 100, replace = TRUE) == 6)
}
```

### Answer

Null hypothesis: <!-- put text after the > below -->

> The die is fair.

Alternative hypothesis: <!-- put text after the > below -->

> The die is not fair.

$p$-value:
```{r}
mean(replicate(10000, p6() >= 23)) * 2
```
(Correct answer: $0.1260$)

Conclusion: <!-- put text after the > below -->

> The die is most likely fair, getting 23 sixes from 100 rolls is not too unusual.


## Self Assessment:

(We're skipping self-assessment this time, since you'll be discussing these in
small groups.)



