---
title: "MTH-235 Assignment 1"
author: "Dennis Ulichney"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: pdf_document
editor_options:
  chunk_output_type: inline
---


This is an R Markdown Notebook. Much like in Colab or Jupyter notebooks, code is
stored in chunks like the one below. You can run it by clicking the play button
in the upper right of the chunk, or by placing your cursor inside the chunk and
selecting "Run Current Chunk" from the drop-down menu under "Run" at the top of
this editor window.

To get started, try running the code block below.

```{r}
plot(cars)
```

All assignments and exams will be submitted through R Markdown Notebooks like
these. For **every** question, your task is to do two things:

1. Write **one line** of code in the answer box that outputs your answer to the
question at hand.

2. If you need more than one line of code to answer the question, put the other
lines of code in the "supporting work" box. See the example below.

---

## Problem 0 (Example)
Evaluate the following expression in R:
$$ \arccos(\cos(2 \pi)) $$

### Supporting Work
```{r}
x <- cos(2 * pi)
```

### Answer
```{r}
acos(x)  # this is the arc cosine function in r
```

Here's a second valid solution to the same problem:

### Supporting Work
```{r}

```

### Answer
```{r}
acos(cos(2 * pi))
```

Since we were able to answer the question with just a single line of code this
time, there was no need to put anything into the Supporting Work block.

---

When you are done with the assignment, your task is to submit TWO artifacts in
Blackboard:

1. this .rmd file
2. the associated .pdf file made by clicking the "Knit" button above

Click the "Knit" button now to preview the file. Then, come back to this editor
and fill in the code blocks below to answer the questions.

**Note that this assignment is intended to be easy.**

---

## Problem 1
Evaluate the fraction $22/7$. (NOTE: Don't worry about the dollar signs; they're
just LaTeX syntax to make the fraction look nice on the PDF output.)

### Supporting Work
```{r}

```

### Answer
```{r}
22/7
```

---

## Problem 2
Find the square root of $500$.

### Supporting Work
```{r}

```

### Answer
```{r}
sqrt(500)
```

---

## Problem 3
Compute $\ln(1,000,000)$.

### Supporting Work
```{r}

```

### Answer
```{r}
log(1000000)
```

