---
title: "MTH-235 Assignment 11"
author: "Dennis Ulichney"
date: '`r gsub("(\\D)0", "\\1", format(Sys.Date(), "%B %d, %Y"))`'
output: pdf_document
editor_options:
  chunk_output_type: inline
---


All assignments and exams will be submitted through R Markdown Notebooks like
these. For **every** question, your task is to do two things:

1. Write **one line** of code in the answer box that outputs your answer to the
question at hand.

2. If you need more than one line of code to answer the question, put the other
lines of code in the "supporting work" box.


```{r}
vgsales <- 
  read.csv("https://raw.githubusercontent.com/optmogames/Video-Games-Sales/master/vgsales.csv",
           header = T)
```


## Problem 1
Run the code block above to import the `vgsales` data frame, which includes data
on video game sales scraped from \url{vgchartz.com}. (For more information on 
this data frame, visit \url{https://www.kaggle.com/gregorut/videogamesales}).

What variables does this data frame include, and what are their types?

### Supporting Work
```{r}
str(vgsales)
```

### Answer

<!-- Short answer: Type your answer after the > -->

> This data frame features the variables rank, name, platform, year, genre, publisher,
North American sales, European Union sales, Japanese sales, sales in other regions, and global sales.
The rank variable is an integer when name, platform, year, genre, and publisher are
all character variables and all of the sales variables are numerical variables.

(Correct answer not disclosed for this problem.)

---

## Problem 2
There are 16,598 observations in the `vgsales` data frame. However, some of the 
observations have the same `Name`; usually, this is because a single game was 
sold on multiple different platforms. How many different titles (as specified by
the `Name`) are there in the data frame?

### Supporting Work
```{r}

```

### Answer
```{r}
length(unique(vgsales$Name))
```

(Correct answer: $11,493$)

---


## Problem 3
How many of the 16,598 video games in the data frame were published by Nintendo?

### Supporting Work
```{r}

```

### Answer
```{r}
sum(vgsales$Publisher == "Nintendo")
```

(Correct answer: $703$)

---

## Problem 4
What percent of the games (as represented by distinct rows) in the data frame 
sold more units in Japan than in Europe?

### Supporting Work
```{r}

```

### Answer
```{r}
mean(vgsales$EU_Sales < vgsales$JP_Sales) * 100
```

(Correct answer: $27.7\%$)

---

## Problem 5
What is the year of publication of the oldest video game in the data frame, 
ignoring the entries with missing dates?

**NOTE**: There's a technical issue you'll have to solve here! First, notice
that the `Year` variable does not have a numeric type; this is because there are
some missing entries which are filled in with the string characters `N/A`, which
results in R rendering the entire variable as a string type. So, you'll need to
do two things:
- coerce the variable to the `numeric` type
- convince R to ignore the missing values
We haven't discussed how to do these things. Use the internet to figure out how
to do these two tasks.


### Supporting Work
```{r}
vgyears <- strtoi(vgsales$Year[vgsales$Year != "N/A"])
```

### Answer
```{r}
min(vgyears)
```

(Correct answer: $1980$)

---

# Interlude

For several of the problems below, you may find the `which.max()` function to be
useful. We did not introduce this explicitly in class, so here's a quick
demonstration:

```{r}
which.max(c(1000, 3000, 6000, 2000, 5000, 2000))
```
It is also possible to solve every problem below by avoiding `which.max()`
entirely. It's your call.


## Problem 6

Run the code block above to import the `vgsales` data frame, which includes data
on video game sales scraped from \url{vgchartz.com}. (For more information on 
this data frame, visit \url{https://www.kaggle.com/gregorut/videogamesales}).

The game "Portal 2" was sold on three different platforms, so it appears as
three separate observations in this data frame (that all have the same name).
On which platform did Portal 2 sell the most units (in terms of global sales)?

### Supporting Work
```{r}
portalsales <- vgsales$Global_Sales[vgsales$Name == "Portal 2"]
portalsystems <- vgsales$Platform[vgsales$Name == "Portal 2"]
```

### Answer
```{r}
portalsystems[which.max(portalsales)]
```

(Answer: `X360`, or XBox 360. Remember, when there's a code block for your
answer, your job is to write a line of code that outputs that answer; in this
case, the output should just be `X360`, not some larger object that contains
that as a sub-part.)

---

## Problem 7
In terms of global sales, what is the all-time best-selling game on the Sega 
Genesis platform (`GEN`)?

### Supporting Work
```{r}
GENsales <- vgsales$Global_Sales[vgsales$Platform == "GEN"]
GENgames <- vgsales$Name[vgsales$Platform == "GEN"]
```

### Answer
```{r}
GENgames[which.max(GENsales)]
```

(Answer: Sonic the Hedgehog 2)

---

## Problem 8
In terms of global sales, what is the all-time best-selling Puzzle-genre game 
for the PlayStation 2 platform (`PS2`)? 

**Note:** I think their categories are kind of terrible. For example, Pac-Man is
clearly *not* a puzzle game even though it is labeled as such. But hey, nobody 
asked me. 

### Supporting Work
```{r}
PS2_Puzzle_Sales <- vgsales$Global_Sales[vgsales$Genre == "Puzzle" & vgsales$Platform == "PS2"]
PS2_Puzzle_Games <- vgsales$Name[vgsales$Genre == "Puzzle" & vgsales$Platform == "PS2"]
```

### Answer
```{r}
PS2_Puzzle_Games[which.max(PS2_Puzzle_Sales)]
```

(Answer: Tetris Worlds)

---

## Problem 9
In terms of *North American* sales, what is the best-selling racing game
published by someone *other* than Nintendo?

### Supporting Work
```{r}
racingsales <- vgsales$NA_Sales[vgsales$Genre == "Racing" & vgsales$Publisher != "Nintendo"]
racinggames <- vgsales$Name[vgsales$Genre == "Racing" & vgsales$Publisher != "Nintendo"]
```

### Answer
```{r}
racinggames[which.max(racingsales)]
```

(Answer: Gran Turismo 3: A-Spec)

---

## Problem 10
How many different observations (that is, rows) in the data frame begin with 
"Yoshi"? *Hint*: Consider the role `substr()` might play here. 

### Supporting Work
```{r}

```

### Answer
```{r}
sum(substr(vgsales$Name, 1, 5) == "Yoshi")
```

(Answer: 11)

---

## Problem 11
In terms of global sales, what is the \emph{worst}-selling SNES-platform game of
all time that has the word "Mario" somewhere in its name?

**Note:** To answer this, you need to be able to search for "Mario" within a
string of characters. The `substr()` function can't do this, since you don't
know where in the title the word "Mario" might appear. We have not encountered
this function in our notebooks or in the Cheat Sheet; your task is to look to 
the internet for help. There are several ways to proceed, at least one of which 
does *not* require you to download an additional library.

### Supporting Work
```{r}
Mariosales <- vgsales$Global_Sales[grepl("Mario", vgsales$Name) & vgsales$Platform == "SNES"]
Mariogames <- vgsales$Name[grepl("Mario", vgsales$Name) & vgsales$Platform == "SNES"]
```

### Answer
```{r}
Mariogames[which.min(Mariosales)]
```

(Answer: Mario no Super Picross)

---

## Problem 12

What is the ratio of total Mario franchise global sales to total Zelda franchise
global sales? (Define the Mario and Zelda franchises as consisting of games that
have either of those words in their titles.)

### Supporting Work
```{r}
Mariototal <- sum(vgsales$Global_Sales[grepl("Mario", vgsales$Name)])
Zeldatotal <- sum(vgsales$Global_Sales[grepl("Zelda", vgsales$Name)])
```

### Answer
```{r}
Mariototal / Zeldatotal
```

(Answer: $6.72$)

---

## Self Assessment:

If this assignment is graded out of 15 points (*note the unusual number*), how 
many points do you think you should earn, **and why?**

<!-- type your response to the self assessment below this line -->
I believe that I should earn all 15 points because I took the time that I needed
to solve all of the problems correctly.

---

## Soliciting Feedback:

This is a genuine question: do you think the self-assessment mechanism on
homework was useful, not useful, or neither? **Why?** Note that your answer will
not be held against you, no matter what it is.

<!-- type your response to the feedback prompt below this line -->
For me, the self-assessment on the homework assignments had no effect. However,
I believe that keeping the self-assessments is a good idea because it provides
a good way for the instructor to keep track of how confident the students are
and what, if anything, they are struggling with.


