---
title: "MTH-235 Assignment 2"
author: "Dennis Ulichney"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: pdf_document
editor_options:
  chunk_output_type: inline
---


All assignments and exams will be submitted through R Markdown Notebooks like
these. For **every** question, your task is to do two things:

1. Write **one line** of code in the answer box that outputs your answer to the
question at hand.

2. If you need more than one line of code to answer the question, put the other
lines of code in the "supporting work" box.

## Problem 1
Find the sum of all integers between 50 and 150 (inclusively).

### Supporting Work
```{r}

```

### Answer
```{r}
sum(50:150)
```

(Correct answer: $10,100$)

---

## Problem 2
Find the sum of the numbers in the sequence 7, 10, 13, ..., 997, 1000.

### Supporting Work
```{r}

```

### Answer
```{r}
sum(seq(7,1000,by=3))
```

(Correct answer: $167,162$)


---

## Problem 3
Write a "deck of cards" vector: `AH, 2H, 3H, ..., TS, JS, QS, KS.` The order of
the cards is not important, but every value (A, 2, 3, 4, 5, 6, 7, 8, 9, T, J, Q,
K) should be matched with every suit (H, D, C, S).

### Supporting Work
```{r}
v1 <- c('A',2:9,'T','J','Q','K')
v2 <- c('H','C','D','S')
```

### Answer
```{r}
paste0(rep(v1,times=4),rep(v2,each=13))
```

(Correct answer: the vector of length 52 described in the problem statement)

---

## Problem 4

Find a decimal approximation of the product
$$
\prod_{n=1}^{1000} \left( \frac{4 n^2}{4n^2 - 1} \right)
= \frac{4 \cdot 1^2}{4 \cdot 1^2 - 1} 
\cdot \frac{4 \cdot 2^2}{4 \cdot 2^2 - 1} 
\dots \frac{4 \cdot (1000)^2}{4 \cdot (1000)^2 - 1}
= \frac{4}{3} \cdot \frac{16}{15}
\cdots \frac{4,000,000}{3,999,999}.
$$

**NOTE**: In RStudio, you can hover over LaTeX syntax like the above to see it
rendered in a human-readable form. It will also appear in a human-readable form
when you knit the PDF output.

**HINT**: Suppose you ask R to divide one vector by another. What do you think
the output *should* be?

### Supporting Work
```{r}

```

### Answer
```{r}
prod((4*(1:1000)^2)/(4*(1:1000)^2-1))
```

(Correct answer: $\approx 1.5704$, which happens to be awfully close to $\pi/2$)

---

## Self Assessment:

If this assignment is graded out of 10 points, how many points do you think you
should earn, **and why?**

<!-- type your response to the self assessment below this line -->
I believe that I should earn the full amount because I completed all of the questions correctly without copying either the work of others or the work that I had done previously.



