---
title: "MTH-235 Assignment 3"
author: "Dennis Ulichney"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: pdf_document
editor_options:
  chunk_output_type: inline
---


All assignments and exams will be submitted through R Markdown Notebooks like
these. For **every** question, your task is to do two things:

1. Write **one line** of code in the answer box that outputs your answer to the
question at hand.

2. If you need more than one line of code to answer the question, put the other
lines of code in the "supporting work" box.


## Problem 1

The game Scrabble contains 98 tiles with letters on them. (Some letters are
repeated, obviously.) In the Supporting Work block below, you will find a vector
representing these tiles. Simulate drawing a collection of 7 Scrabble tiles 
(without replacement).

### Supporting Work
```{r}
scrabble <- rep(LETTERS, c(9, 2, 2, 4, 12, 2, 3, 2, 9, 1,
                           1, 4, 2, 6, 8, 2, 1, 6, 4, 6,
                           4, 2, 2, 1, 2, 1))  # makes a vector of length 98
```

### Answer
```{r}
sample(scrabble,7)
```

(Correct answer: a collection of 7 random letters -- be sure you have correctly
specified the `replace` parameter!)

---

## Problem 2
In the game Bingo, a dealer draws several balls from a bin. There are 75 balls
in the bin, each labeled with both a letter and a number. The labels are B-1, 
B-2, ..., B-15, I-16, I-17, ..., I-30, N-31, N-32, ..., N-45, G-46, G-47, ...,
G-60, O-61, O-62, ..., O-75.

Simulate drawing a collection of ten of these balls randomly (without replacement)
from the bin.

### Supporting Work
```{r}
BINGO <- paste(rep(c("B","I","N","G","O"),each=15),(1:75),sep="-")
```

### Answer
```{r}
sample(BINGO,10)
```

(Correct answer: A random sequence of 10 things that look like `N-50` or `B-7`,
complete with the hyphens.)

---

## Problem 3
Simulate drawing 5 cards (without replacement) from a standard deck of 52.
Recall that you constructed the "deck of cards" vector in the previous
assignment.

### Supporting Work
```{r}
Deck <- paste0(rep(c('A',2:9,'T','J','Q','K'),times=4),rep(c('H','C','D','S'),each=13))
```

### Answer
```{r}
sample(Deck,5)
```

(Correct answer: a list of five random cards such as "5H", "8S", or "QC")

---

## Problem 4
Background: a vector can be given additional structure with the use of the
`names()` function; see the example below in the Supporting Work section.

A child chooses her own outfit. For each of her shirt, pants, and socks, she
will select something that is either red, green, or blue. She may select the
same color for multiple components of her outfit. She is twice as likely
to select red as she is to select green, and she is three times as likely to
select red as she is to select blue. 

The `outfit` vector below is a static (i.e. non-random) vector; run the answer
block to see its structure. Then, overwrite it so that the components of the 
vector are random chosen to be red, green, or blue according to the 
probabilities described above.

### Supporting Work
```{r}
outfit <- c(sample(c("red","green","blue"),3,prob=c(6,3,2)))           # static vector of colors
names(outfit) <- c("shirt", "pants", "socks")   # give the vector elements names
```

### Answer
```{r}
outfit  # don't change this block; change the supporting work block instead
```

(Correct answer: a named vector with names "shirt", "pants", and "socks" as 
above, but whose entries are randomly populated with red, green, and blue. Make 
sure the colors are chosen according to the probabilities described in the 
problem statement!)

---

## Problem 5
The `temperatures` vector, defined below, consists of temperature readings of
731 human patients. A "normal" human temperature is defined as one that is
between 97.7 and 99.5 degrees. What proportion of the temperatures in the
vector qualify as normal? (Note: none of the temperatures are exactly 97.7 or
99.5.)

### Supporting Work
```{r}
# (run this block, but otherwise leave it alone)
set.seed(1845)
temperatures <- rnorm(731, 98.6, 0.5) # don't worry about rnorm() yet
set.seed(as.numeric(Sys.time()))
```

### Answer
```{r}
mean((temperatures > 97.7) & (temperatures < 99.5))
```

(Correct answer: $0.9357$, or $93.57\%$)

---

## Problem 6
Two people are playing a game. Player 1 rolls a D6; player 2 rolls a D4. They
each roll their respective dice 100 times. The higher number wins the round;
since Player 2 starts at a disadvantage, she also is declared the winner in the
case of a tie.

The records of their games are stored in the `player1` and `player2` vectors in
the Supporting Work block below. How many games did Player 2 win?

### Supporting Work
```{r}
# (run this block, but otherwise leave it alone)
set.seed(999)
player1 <- sample(1:6, 100, replace = T)
player2 <- sample(1:4, 100, replace = T)
set.seed(as.numeric(Sys.time()))
```

### Answer
```{r}
sum((player2 > player1) | (player2 == player1))
```

(Correct answer: $42$)

---

## Problem 7

Someone flips a pair of coins many times. The results of these flips are stored
in the `coins` vector below. Each entry of the vector corresponds to a *pair* of
flips, so it looks like either `HH`, `HT`, `TH`, or `TT`.

In what proportion of the flips did the player get two of the same outcome?

### Supporting Work
```{r}
# (run this block, but otherwise leave it alone)
set.seed(1024)
coins <- sample(c("HH", "HT", "TH", "TT"), 500, replace = T)
set.seed(as.numeric(Sys.time()))
```

### Answer
```{r}
mean((coins == "HH") | (coins == "TT"))
```

(Correct answer: $0.502$, or $50.2\%$)

---

## Self Assessment:

If this assignment is graded out of 10 points, how many points do you think you
should earn, **and why?**

<!-- type your response to the self assessment below this line -->

  I believe that I should earn the full amount of points because my work is correct and complete.



