---
title: "MTH-235 Assignment 4"
author: "Dennis Ulichney"
date: '`r gsub("(\\D)0", "\\1", format(Sys.Date(), "%B %d, %Y"))`'
output: pdf_document
editor_options:
  chunk_output_type: inline
---


All assignments and exams will be submitted through R Markdown Notebooks like
these. For **every** question, your task is to do two things:

1. Write **one line** of code in the answer box that outputs your answer to the
question at hand.

2. If you need more than one line of code to answer the question, put the other
lines of code in the "supporting work" box.

## Problem 1
Write a function, `c_to_f()`, that takes a (numerical) input as a temperature in
degrees Celsius and converts it to a temperature in degrees Fahrenheit. See
[here](https://en.wikipedia.org/wiki/Celsius#See_also) for a reference of the
underlying mathematics.

### Supporting Work
```{r}
c_to_f <- function(x){
  # fill this part in
  (9/5)*x + 32
}
```

### Answer
```{r}
# leave this entire code block alone
c_to_f(c(0, 37, 100))
```

(Correct answers: $32$, $98.6$, $212$)

---

## Problem 2
Write a function that takes a numeric *vector* as input and produces as an
output the *range* of the vector, which is defined as the distance between its
largest and smallest value. Test your function on the `test_vector_1`, which is
defined in the Supporting Work block below. 

You should write your function in the Supporting Work block and test it in the
Answer block; see the previous problem for an example.

**NOTE**: You should not name your function `range()`, because there is already 
an R function with that name that does something else.


### Supporting Work
```{r}
# ===== leave code below alone =====
set.seed(5)
test_vector_1 <- rnorm(10000)
  # generates 10,000 values with the standard normal distribution
  # we will encounter this concept later in the class
test_vector_2 <- test_vector_1 + rnorm(10000, mean = 0, sd = 0.5)
  # create a second vector to use for a later problem
set.seed(as.numeric(Sys.time()))    # undoes effect of setting random seed
# ===== leave code above alone =====
# put your code below here
range <- function(x){
  max(x) - min(x)
}

```

### Answer
```{r}
range(test_vector_1)
```

(Correct answer: $7.753$)

---

## Problem 3

Suppose $x = (x_1, x_2, \dots, x_n)$ and $y = (y_1, y_2, \dots, y_n)$ are two
numeric vectors of the same length. The **linear correlation coefficient** 
between $x$ and $y$ is defined to be
$$ r = \frac{\sum \left[ (x_i - \overline x)(y_i - \overline y) \right]}
{(n-1) s_x s_y}$$
where $\overline x, \overline y$ are the respective means of $x, y$ and
$s_x, s_y$ are the respective standard deviations of $x, y$. (Recall that the
function in R to compute standard deviations is `sd()`.)

Write a function that accepts as input two numeric vectors of equal length and
produces as output their correlation coefficient. Test your function on the
vectors `test_vector_1` and `test_vector_2`, which were defined in a code block 
above.

**NOTE:** This function is already implemented in R as `cor()`, so you should
name your function something else. Your goal is to make a function and produce
the same output as `cor(test_vector_1, test_vector_2)`.

### Supporting Work
```{r}
CC <- function(x,y){
    sum((x-mean(x))*(y-mean(y)))/((length(x)-1)*sd(x)*sd(y))
}
```

### Answer
```{r}
CC(test_vector_1,test_vector_2)
```

(Correct answer: $0.8941$)

---

## Problem 4
Write a function that takes a positive integer `n` as input, flips $n$ coins, 
and outputs `TRUE` if the coins all land on the same face (and `FALSE` 
otherwise). Test your function in the answer block by supplying the input
$n = 3$. Remember: no `if` structures allowed!

### Supporting Work
```{r}
# write your function here
flip <- function(n){
  length(unique(sample(0:1,n,replace=T))) == 1
}

```

### Answer
```{r}
# test your function with n = 3 here
flip(3)
```

(Correct answer: Your function should output `TRUE` about $1/4$ of the time and
`FALSE` about $3/4$ of the time if evaluated repeatedly.)

---

## Problem 5
A bin contains five components from supplier A, three from supplier B, and two 
from supplier C. Write a function that simulates drawing three components 
without replacement from this bin for testing and counts the number of different
suppliers represented in the sample. (Example: if the components drawn are from
suppliers A, C, A, then the output of the function should be 2, because two
different suppliers are represented.)

**HINT**: One very useful function that we have not previously encountered is
`unique(x)`, which outputs a vector of the unique entries in the vector `x` --
that is, it removes repetitions from `x`. How might you combine this vector with
other functions we have encountered?

### Supporting Work
```{r}
bin <- function(){
  length(unique(sample(rep(c('A','B','C'),times=c(5,3,2)),3)))
}
```

### Answer
```{r}
bin()
```

(Correct answer: Your function should output one of the numbers 1, 2, or 3, and 
the output should change upon repeated evaluations. The most common output
should be 2; the output of 1 is rare.)

---

## Interlude

```{r, include = FALSE}
  # do not edit this block
  # note that the option include = FALSE hides this from your PDF output
cards <- 
  paste0(c("A", "2", "3", "4", "5", "6", "7", "8", "9", "T", "J", "Q", "K"),
         rep(c("H", "C", "S", "D"), each = 13))
  # format of cards is: [value][suit], such as "3H" for 3 of hearts,
  # "7C" for 7 of clubs, "TD" for 10 of diamonds
```

The `cards` vector, which you can place in your environment by running the block
above, contains all cards in a standard 52-card deck. (You have written a
version of this vector in previous assignments.)

To take a card from the `cards` vector and extract either the suit or the value,
you might find the `substr()` function to be useful; refer to the cheat sheet
for syntax and examples. 

---

## Problem 6
Write a function that simulates drawing `n` cards without replacement from the
`cards` vector and counts how many hearts (suit of H) are drawn. Test your
simulation in the answer block with $n = 10$.

### Supporting Work
```{r}
cards <- paste0(c("A", "2", "3", "4", "5", "6", "7", "8", "9", "T", "J", "Q", "K"), rep(c("H", "C", "S", "D"), each = 13))
hearts <- function(n){
  sum(substr(sample(cards,n),2,2) == "H")
}
```

### Answer
```{r}
hearts(10)
```

(Correct answer: your function should output a number between 0 and 10
inclusively upon repeated evaluation. Note that some outputs, such as those
between 6 and 10, are theoretically possible but will be extraordinarily rare in
practice.)

---

## Problem 7
Write a function that simulates drawing `n` cards without replacement from the
`cards` vector and counts how many face cards (value of J, Q, or K) are drawn.
Test your simulation in the answer block with $n = 10$.

### Supporting Work
```{r}
cards <- paste0(c("A", "2", "3", "4", "5", "6", "7", "8", "9", "T", "J", "Q", "K"), rep(c("H", "C", "S", "D"), each = 13))
face <- function(n){
  sum(substr(sample(cards,n),1,1) %in% c("J","Q","K"))
}
```

### Answer
```{r}
face(10)
```

(Correct answer: your function should output a number between 0 and 10 
inclusively upon repeated evaluation. Note that some outputs, such as those
between 6 and 10, are theoretically possible but will be extraordinarily rare in
practice.)

---

## Self Assessment:

If this assignment is graded out of 10 points, how many points do you think you
should earn, **and why?**

<!-- type your response to the self assessment below this line -->
I believe that I should earn the full amount because I put forth enough effort to complete each question correctly.



