---
title: "MTH-235 Assignment 5"
author: "Dennis Ulichney"
date: '`r gsub("(\\D)0", "\\1", format(Sys.Date(), "%B %d, %Y"))`'
output: pdf_document
editor_options:
  chunk_output_type: inline
---


All assignments and exams will be submitted through R Markdown Notebooks like
these. For **every** question, your task is to do two things:

1. Write **one line** of code in the answer box that outputs your answer to the
question at hand.

2. If you need more than one line of code to answer the question, put the other
lines of code in the "supporting work" box.


## Problem 1

If you flip 100 coins, what is the probability that **exactly** 50 of them land 
on heads? (Write a function and replicate it 10,000 times to investigate; this
will be our standard procedure for investigating probabilities.)

### Supporting Work
```{r}
flipsum <- function(){
  sum(sample(c("H","T"),100,replace=T) == "H")
}
```

### Answer
```{r}
mean(replicate(10000,flipsum() == 50))
```

(Correct answer: $\approx 0.0796$; remember that from here on out, your answers
are expected to be close, but probably not exactly equal to, the correct ones.)

---

## Problem 2

If you flip 100 coins, what is the probability that between 40 and 60 of them
(inclusively) land on heads?

### Supporting Work
```{r}
flipsum <- function(){
  sum(sample(c("H","T"),100,replace=T) == "H")
}
```

### Answer
```{r}
mean(replicate(10000,flipsum() >= 40 | flipsum() <= 60))
```

(Correct answer: $\approx 0.9648$)

---

## Problem 3
A bin contains five components from supplier A, three from supplier B, and two 
from supplier C. If three components are randomly selected (without replacement)
for testing, how many different suppliers are represented in the sample? Make a 
histogram of 10,000 simulations; be sure to consider the `breaks`. (Recall: you 
saw this problem setup in the last assignment.)

### Supporting Work
```{r}
bin <- rep(c("A","B","C"),times=c(5,3,2))
supcount <- function(){
  length(unique(sample(bin,3)))
}
```

### Answer
```{r}
hist(replicate(10000,supcount()),breaks = .5 : 3.5)
```

(Correct answer: a histogram with a tall bar centered around 2, a medium-ish bar
centered around 3, and a short bar centered around 1. The bars should be
adjacent to one another with no whitespace between them.)

---

## Problem 4

A bin contains five components from supplier A, three from supplier B, and two 
from supplier C. If three of the components are randomly selected for testing, 
what is the probability that each supplier would have at least one component 
tested? 

### Supporting Work
```{r}
bin <- rep(c("A","B","C"),times=c(5,3,2))
supcount <- function(){
  length(unique(sample(bin,3)))
}
```

### Answer
```{r}
mean(replicate(10000,supcount()==3))
```

(Correct answer: $1/4$)

---

## Problem 5

Suppose you draw `LETTERS` without replacement from a jar. What is the expected
number of draws required to collect all three of the letters A, B, and C?

### Supporting Work
```{r}
drawABC <- function(){
  draw <- sample(LETTERS)
  max(c(which(draw=="A"),which(draw=="B"),which(draw=="C")))
}
```

### Answer
```{r}
mean(replicate(10000,drawABC()))
```

(Correct answer: $\approx 20.25$ draws)

---

## Problem 6

A player in a game must roll either a D4 (a 4-sided die) or a D6 (a 6-sided
die), depending on the outcome of a coin flip. What is their expected die roll?

**HINT:** Inside your function, make two calls to the `sample()` function -- one
corresponding to the coin flip, and the other to the die roll. Use the output of
one of the `sample()` functions as an input to the other.

### Supporting Work
```{r}
fliproll <- function(){
  flip <- sample(c(4,6),1,replace=T)
  roll <- sample(1:flip,1,replace=T)
}
```

### Answer
```{r}
mean(replicate(10000,fliproll()))
```

(Correct answer: $3$)

---

## Problem 7

A person who frequently travels from Cleveland to Chicago drives her car with 
$40\%$ probability and takes the train with $60\%$ probability. If she takes her
car, she has a $50\%$ chance of a 10-minute traffic delay (and a $50\%$ of no 
delay at all); if she takes the train, she has a $10\%$ chance of a one-hour 
traffic delay (and a $90\%$ chance of no delay at all). What is her expected 
delay?

**HINT:** This is a slightly more challenging multistage problem. The problem
would certainly be more straightforward if you were permitted to use `if`
structures -- but that is still off the table, for now. One way to approach this
is to simulate *both* a car delay and a train delay independently, then use the 
`sample()` function again to select one of those.

### Supporting Work
```{r}
delaycalc <- function(){
  car <- sample(c(0,10),1)
  train <- sample(c(0,60),1,prob=c(9,1))
  sample(c(car,train),1,prob=c(4,6))
}
```

### Answer
```{r}
mean(replicate(10000,delaycalc()))
```

(Correct answer: $5.6$ minutes)

---

## Self Assessment:

If this assignment is graded out of 10 points, how many points do you think you
should earn, **and why?**

<!-- type your response to the self assessment below this line -->
I believe that I should earn full credit because I put forth the effort necessary to answer all questions correctly.



