---
title: "MTH-235 Assignment 8"
author: "Dennis Ulichney"
date: '`r gsub("(\\D)0", "\\1", format(Sys.Date(), "%B %d, %Y"))`'
output: pdf_document
editor_options:
  chunk_output_type: inline
---


All assignments and exams will be submitted through R Markdown Notebooks like
these. For **every** question, your task is to do two things:

1. Write **one line** of code in the answer box that outputs your answer to the
question at hand.

2. If you need more than one line of code to answer the question, put the other
lines of code in the "supporting work" box.

---

## Problem 1
A gambler makes a sequence of bets. Each bet is a $\$1$ bet in which she has a
$50\%$ chance of winning $\$1$ or losing $\$1$. She begins with a $\$2$ bankroll
and continues betting until she either loses all her money or has a $\$10$ 
bankroll. What is the probability that she will leave with $\$10$ (instead of 
leaving with nothing)?

### Supporting Work
```{r}
p1 <- function(){
  bankroll <- 2
  while (bankroll > 0 & bankroll < 10){
    win <- sample(0:1,1)
    if (win == 1){
      bankroll <- bankroll + 1
    }
    else {
      bankroll <- bankroll - 1
    }
  }
  bankroll == 10
}
```

### Answer
```{r}
mean(replicate(10000, p1()))
```

(Correct answer: $20\%$)

---

## Problem 2
In the MLB World Series, teams play against each other until one team wins 
4 games. Suppose Cleveland has a $45\%$ chance of beating Chicago in each game. 
What is the probability that Cleveland wins the world series?

### Supporting Work
```{r}
p2 <- function(){
  Cleveland <- 0
  Chicago <- 0
  while(Cleveland < 4 & Chicago < 4){
     Clevewin <- sample(0:1, 1, prob=c(.55, .45))
     if(Clevewin == 1){
       Cleveland <- Cleveland + 1
     }
     else {
       Chicago <- Chicago + 1
     }
  }
  Cleveland == 4
}
```

### Answer
```{r}
mean(replicate(10000, p2()))
```

(Correct answer: $\approx 0.3917$)

---

## Problem 3

Suppose you repeatedly draw upper-case letters from a jar without replacement 
until the first time you get one that is out of order. What is the expected 
number of draws for this to occur? (Example: if you draw B-E-R-D, you would stop
at D, because it was the first letter you drew that didn't occur after the 
previous letter in the alphabet.)	

**HINT**: To make your life easier, on each experiment you might use
`sample(LETTERS)` to draw a full sample of all 26 letters in a random order and
work from there.

**Technical note**: It is OK to ignore the case where you draw all the letters
in perfect order, because this is rare enough that it will probably not occur 
even in trillions of simulations.

### Supporting Work

```{r}
p3 <- function(){
  draw <- sample(1:26)
  count <- 2
  while(draw[count] < draw[count-1]){
    count <- count + 1
  }
  count
}
```

### Answer

```{r}
mean(replicate(10000, p3()))
```

(Correct answer: about $2.72$ draws)

---

## Problem 4

A COVID-19 serology test developed by Cellex Inc. tests the blood of patients to
determine whether they have previously carried the coronavirus. The 
**sensitivity** of the test, or the probability that they test positive given 
that they had the virus, is $93.8\%$. Similarly, the **specificity** of the 
test, or the probability that they test negative given that they have not had 
the virus, is $95.6\%$. (Note: All information in this paragraph is
[true](https://www.centerforhealthsecurity.org/resources/COVID-19/serology/Serology-based-tests-for-COVID-19.html).) 

In a certain region, about $10\%$ of residents have been exposed to the 
coronavirus. What is the probability that a randomly-selected individual has had
the disease, given that they have a positive test? 

### Supporting Work

```{r}
p4 <- function(){
  exposure <- sample(0:1, 1, prob = c(90, 10))
  if(exposure == 1){
    positive <- sample(0:1, 1, prob = c(6.2, 93.8))
  }
  else {
    positive <- sample(0:1, 1, prob = c(95.6, 4.4))
  }
  (exposure)[positive]
}
```

### Answer

```{r}
mean(unlist(replicate(10000, p4())))
```

(Correct answer: $\approx 0.7031$)

---

## Problem 5

A plane seats 100 passengers. The first passenger to board a full flight loses
his ticket, so he randomly selects one of the available seats. Every subsequent
passenger sits in their assigned seat (if open) or picks one of the available
seats randomly if not. What is the probability that the final passenger will sit
in her own seat? 

(Correct answer: exactly $50\%$, and if you're anything like me, that's a
surprising answer)

### Note

This problem can be completed using only things explicitly found in the Cheat
Sheet up through section 3.5. However, you are strongly encouraged to hunt on
the internet for new functions that we haven't used if you think they would 
serve you well for this problem. (The only thing you *shouldn't* do is post this
exact question on Stack Overflow or something like that.)

### Supporting Work

```{r}
# The code here was copied from my solution to the same problem in Worksheet 3.5.
p5 <- function(){
  p1 <- sample(1:100,1)
  seats <- (1:100)[-p1]
  for (p in 1:98){
    if (p %in% seats){
      seats <- seats[-which(p==seats)]
    }

    else {
      seats <- seats[-which(sample(seats,1)==seats)]
    }
  }
  100 %in% seats
}
```

### Answer

```{r}
mean(replicate(10000, p5()))
```

---

## Self Assessment:

If this assignment is graded out of 10 points, how many points do you think you
should earn, **and why?**

<!-- type your response to the self assessment below this line -->
I believe that I should earn the full amount because I took the time that I needed to answer each question correctly.



