---
title: "MTH-235 Assignment 9"
author: "Dennis Ulichney"
date: '`r gsub("(\\D)0", "\\1", format(Sys.Date(), "%B %d, %Y"))`'
output: pdf_document
editor_options:
  chunk_output_type: inline
---


All assignments and exams will be submitted through R Markdown Notebooks like
these. For **every** question, your task is to do two things:

1. Write **one line** of code in the answer box that outputs your answer to the
question at hand.

2. If you need more than one line of code to answer the question, put the other
lines of code in the "supporting work" box.

## Problem 1
Upon studying low bids for shipping contracts, a microcomputer manufacturing
company finds that intrastate contracts have low bids that are uniformly
distributed between $\$20,000$ and $\$25,000$. Find the probability that the low
bid on the next intrastate shipping contract is below $\$22,000$, given that it
is below $\$23,000$.


### Supporting Work
```{r}
p1 <- function(){
  bid <- runif(1, 20000, 25000)
  (bid < 22000)[bid < 23000]
}
```

### Answer
```{r}
mean(unlist(replicate(10000, p1())))
```

(Correct answer: $2/3$)

---

## Problem 2

An ice-maker makes ice cubes of random size. Each cube has a side length that is
uniformly distributed on $[1, 2]$. (All sides must have the same length since
this is a cube.) What is the expected side length of such a cube?

### Supporting Work
```{r}

```

### Answer
```{r}
mean(replicate(10000, runif(1, 1, 2)))
```

(Correct answer: $1.5$, and you should stop and think about why that number
makes perfect sense in the context of this problem.)

---

## Problem 3
An ice-maker makes ice cubes of random size. Each cube has a side length that is
uniformly distributed on $[1, 2]$. (All sides must have the same length since
this is a cube.) What is the expected **volume** of such a cube?

### Supporting Work
```{r}

```

### Answer
```{r}
mean(replicate(10000, runif(1, 1, 2)**3))
```

(Correct answer: $15/4 = 3.75$. Note in particular that although the expected
side length was $1.5$, the expected volume is *not* $1.5^3 = 3.375$.)

---

## Problem 4
Two people are positioned randomly in a 10' by 10' room. For social distancing
purposes, they would like to remain at least 6' apart from one another. What is
the probability that this has occurred?

**Hint 1:** The position of one person can be modeled as having coordinates
$(X, Y)$, each of which are uniformly distributed on $[0, 10]$.

**Hint 2:** Remember the distance formula from some previous math class -- or, 
if you have forgotten it, look it up.

### Supporting Work
```{r}
p4 <- function(){
  pos1 <- c(runif(1, 0, 10), runif(1, 0, 10))
  pos2 <- c(runif(1, 0, 10), runif(1, 0, 10))
  ((pos1[1] - pos2[1])^2 + (pos1[2] - pos2[2])^2)^.5 >= 6
}
```

### Answer
```{r}
mean(replicate(10000, p4()))
```

(Correct answer: around $38\%$)

---

## Problem 5
A company that manufactures and bottles apple juice uses a machine that
automatically fills 16-oz bottles. There is some variation, however, in the
amounts of liquid dispensed in the bottles that are filled. The amount dispensed
has been observed to be approximately normally distributed with mean 16.0 oz
and standard deviation 1.0 oz. What is the probability that one bottle will have
more than 17.0 oz dispensed into it?

### Supporting Work
```{r}

```

### Answer
```{r}
mean(replicate(10000, rnorm(1, 16, 1) > 17))
```

(Correct answer: $0.1587$)

---

## Problem 6
Refer to the previous problem. What is the probability that a bottle will be
filled with more than 14.5 oz, given that it is filled with less than 16.5 oz?

### Supporting Work
```{r}
p6 <- function(){
  fill <- rnorm(1, 16, 1)
  (fill > 14.5)[fill < 16.5]
}
```

### Answer
```{r}
mean(unlist(replicate(10000, p6())))
```

(Correct answer: $0.9034$)

---

## Problem 7
Refer to the previous two problems. If three bottles are filled, what is the
probability that all three bottles will have between 14.5 and 16.5 oz?

### Supporting Work
```{r}
p7 <- function(){
  fill3 <- c(rnorm(1, 16, 1), rnorm(1, 16, 1), rnorm(1, 16, 1))
  length(fill3[fill3 >= 14.5 & fill3 <= 16.5]) == 3
}
```

### Answer
```{r}
mean(replicate(10000, p7()))
```

(Correct answer: $0.2437$)

---

## Problem 8
The length of time $Y$ needed to complete a key operation in the construction of
houses has a normal distribution with $\mu = 10$ and $\sigma = 2$ (both in
hours). The cost (in dollars) of completing the operation is found by the 
formula $C = 100 + 40Y + 3Y^2$. What is the expected cost of completing the
operation?

### Supporting Work
```{r}
p8 <- function(){
  time <- rnorm(1, 10, 2)
  3*(time^2) + 40*(time) + 100
}
```

### Answer
```{r}
mean(replicate(10000, p8()))
```

(Correct answer: $812$)

---

## Self Assessment:

If this assignment is graded out of 10 points, how many points do you think you
should earn, **and why?**

<!-- type your response to the self assessment below this line -->
I believe that I should earn full credit because I took the time necessary to
complete all of the problems correctly.



